---
title: "RTLib Validation"
output: rtlib_validation
---

## Init
Load libraries, common functions
Load datasets
 * ev.pep: Peptide-centric, Fit2 STAN (Updated 11/18/2017)
 * ev.exp: Experiment-centric, Fit2 STAN (Updated 11/18/2017)
 * ev.pwise: Pairwise

```{r}
library(tidyverse)
source('lib.R')

ev.pep <- read_tsv('dat/ev.adj.Fit2.txt')
ev.exp <- read_tsv('dat/ev.adj.Fit2.expcentric.txt')
ev.pwise <- read_tsv('dat/evidence+dRT.elite.txt')
```

## Basic evaluation of validation algorithm, on peptide-centric adjustment on experiment 19A

 * _Original_ Distribution: Original PEP < 0.05
 * _New_ Distribution: New PEP (PEP.new) < 0.05, and Original PEP > 0.05 (newly upgraded PSMs). Should be disjoint from _Original_ Distribution
 * _Null_ Distribution: Fake proteins assembled from randomly matched peptides
 * _Total_ Distribution (Not shown): New PEP < 0.05. *Not* the union between _Original_ and _New_, as some PSMs with Original PEP < 0.05 could be downgraded so that PEP.new > 0.05.

```{r}
cors <- validate.lib(ev.pep, exps=c('19A'))
cors %>%
  filter(Type != 'Total') %>%
  ggplot(aes(x=data, color=Type)) +
  geom_line(stat='density', position='identity')

```


## Compare performance with bulk J + H dataset, 19A
19A is comprised of several raw files/runs, some targeted

```{r}
cors.exp <- validate.lib(ev.exp, exps=c('19A'))
cors.pep <- validate.lib(ev.pep, exps=c('19A'))
cors.pwise <- validate.lib(ev.pwise, ev.type='Ali', exps=c('19A'))

cors.exp$Method <- 'STAN - Experiment-Centric'
cors.pep$Method <- 'STAN - Peptide-Centric'
cors.pwise$Method <- 'Pairwise'
cors.exp$Experiment <- '19A'
cors.pep$Experiment <- '19A'
cors.pwise$Experiment <- '19A'

cors <- rbind(cors.exp, cors.pep, cors.pwise)


#cors.pwise %>%
cors %>%
  filter(!Type %in% c('Total')) %>%
ggplot(aes(x=data, color=Type)) +
  geom_line(stat='density', position='identity') +
  facet_grid(~Method) +
  scale_color_manual(labels=c('New: PEP.new < 0.05', 'Null', 'Original: PEP < 0.05'), 
                     values=c('green', 'red', 'blue')) +
  labs(title='RTLib Validation: 19A', x='Correlation')
```


```{r}
cors.exp.a <- validate.lib(ev.exp, exps=c('30'))
cors.pep.a <- validate.lib(ev.pep, exps=c('30'))
cors.pwise.a <- validate.lib(ev.pwise, ev.type='Ali', exps=c('30'))

cors.exp.a$Method <- 'STAN - Experiment-Centric'
cors.pep.a$Method <- 'STAN - Peptide-Centric'
cors.pwise.a$Method <- 'Pairwise'
cors.exp.a$Experiment <- '30'
cors.pep.a$Experiment <- '30'
cors.pwise.a$Experiment <- '30'

cors <- rbind(cors, cors.exp.a, cors.pep.a, cors.pwise.a)

cors %>%
  filter(!Type %in% c('Total')) %>%
ggplot(aes(x=data, color=Type)) +
  geom_line(stat='density', position='identity') +
  facet_grid(Experiment~Method) +
  scale_color_manual(labels=c('New: PEP.new < 0.05', 'Null', 'Original: PEP < 0.05'), 
                     values=c('green', 'red', 'blue')) +
  labs(title='RTLib Validation: 19A', x='Correlation')
```